\documentclass{article}
\usepackage{amsmath}
\usepackage{circuitikz}
\usepackage{enumitem}

\begin{document}
	
	\section*{Comprehensive Electronics \& Arduino Quiz}
	
	\begin{enumerate}
		\item What is the primary function of a capacitor in a DC circuit?
		\begin{enumerate}
			\item To amplify current
			\item To store electrical charge
			\item To generate voltage
			\item To measure resistance
		\end{enumerate}
		\textbf{Answer: b) To store electrical charge} \\
		\textbf{Explanation:} Capacitors store and release electrical energy, smoothing voltage fluctuations in DC circuits.
		
		\item The relationship between voltage (V), current (I), and resistance (R) is described by:
		\begin{enumerate}
			\item $P = IV$
			\item $V = \frac{I}{R}$
			\item $V = IR$
			\item $R = \frac{V}{I^2}$
		\end{enumerate}
		\textbf{Answer: c) $V = IR$} \\
		\textbf{Explanation:} Ohm's Law states voltage equals current multiplied by resistance ($V = IR$).
		
		\item What does PWM stand for in Arduino context?
		\begin{enumerate}
			\item Pulse Width Modulation
			\item Power Wave Management
			\item Programmable Wave Model
			\item Pulse Wave Measurement
		\end{enumerate}
		\textbf{Answer: a) Pulse Width Modulation} \\
		\textbf{Explanation:} PWM controls power by rapidly switching between HIGH and LOW states.
		
		\item The typical operating voltage for most Arduino boards is:
		\begin{enumerate}
			\item 12V
			\item 9V
			\item 3.3V
			\item 5V
		\end{enumerate}
		\textbf{Answer: d) 5V} \\
		\textbf{Explanation:} Most Arduino I/O pins operate at 5V, though some boards have 3.3V logic.
		
		\item In PCB design, what is the purpose of a via?
		\begin{enumerate}
			\item To connect traces on different layers
			\item To mount surface-mount components
			\item To create ground planes
			\item To insulate copper traces
		\end{enumerate}
		\textbf{Answer: a) To connect traces on different layers} \\
		\textbf{Explanation:} Vias are conductive holes that connect copper layers in multilayer PCBs.
		
		\item What will this Arduino code output to the serial monitor?
		\begin{verbatim}
			void setup() {
				Serial.begin(9600);
				Serial.println(3 + 2 * 4);
			}
			void loop() {}
		\end{verbatim}
		\begin{enumerate}
			\item 20
			\item 11
			\item 14
			\item Error
		\end{enumerate}
		\textbf{Answer: b) 11} \\
		\textbf{Explanation:} Multiplication has higher precedence than addition (2*4=8, 3+8=11).
		
		\item The purpose of a pull-up resistor is to:
		\begin{enumerate}
			\item Limit current to an LED
			\item Ensure defined voltage when switch is open
			\item Amplify sensor signals
			\item Convert analog to digital signals
		\end{enumerate}
		\textbf{Answer: b) Ensure defined voltage when switch is open} \\
		\textbf{Explanation:} Pull-up resistors prevent floating inputs by providing default HIGH state.
		
		\item What does the "D" in "DIP" package stand for?
		\begin{enumerate}
			\item Digital
			\item Dual
			\item Direct
			\\item Discrete
		\end{enumerate}
		\textbf{Answer: b) Dual} \\
		\textbf{Explanation:} DIP = Dual In-line Package with two parallel rows of pins.
		
		\item Which component is used for overvoltage protection?
		\begin{enumerate}
			\item Capacitor
			\item Varistor
			\item Thermistor
			\item Potentiometer
		\end{enumerate}
		\textbf{Answer: b) Varistor} \\
		\textbf{Explanation:} Varistors (MOVs) clamp voltage spikes by changing resistance.
		
		\item What is the primary function of a voltage regulator?
		\begin{enumerate}
			\item To store electrical energy
			\item To maintain constant output voltage
			\item To amplify current
			\item To generate clock signals
		\end{enumerate}
		\textbf{Answer: b) To maintain constant output voltage} \\
		\textbf{Explanation:} Regulators maintain stable voltage despite input variations or load changes.
		
		\item In Arduino, how many analog input pins are on a standard Uno board?
		\begin{enumerate}
			\item 4
			\item 6
			\item 8
			\item 10
		\end{enumerate}
		\textbf{Answer: b) 6} \\
		\textbf{Explanation:} Arduino Uno has analog inputs A0-A5 (6 channels).
		
		\item What is the standard trace width for signal traces in hobbyist PCBs?
		\begin{enumerate}
			\item 0.1mm
			\item 0.3mm
			\item 0.8mm
			\item 1.5mm
		\end{enumerate}
		\textbf{Answer: b) 0.3mm} \\
		\textbf{Explanation:} 0.3mm (12mil) is common for signal traces carrying <500mA.
		
		\item Which function initializes a digital pin as output?
		\begin{enumerate}
			\item pinMode(pin, OUTPUT)
			\item digitalWrite(pin, HIGH)
			\item setPin(pin, OUT)
			\item analogWrite(pin, 255)
		\end{enumerate}
		\textbf{Answer: a) pinMode(pin, OUTPUT)} \\
		\textbf{Explanation:} pinMode() configures pin direction before use.
		
		\item What is solder mask's primary purpose on PCBs?
		\begin{enumerate}
			\item Increase conductivity
			\item Prevent oxidation
			\item Provide component labels
			\item Insulate copper traces
		\end{enumerate}
		\textbf{Answer: d) Insulate copper traces} \\
		\textbf{Explanation:} Solder mask prevents accidental shorts and protects copper.
		
		\item What does this Arduino code do?
		\begin{verbatim}
			void setup() {
				DDRD = 0xFF;
			}
			void loop() {
				PORTD = 0xAA;
			}
		\end{verbatim}
		\begin{enumerate}
			\item Reads analog inputs
			\item Sets all digital pins 0-7 as outputs with alternating HIGH/LOW
			\item Activates PWM on pin 9
			\item Measures temperature
		\end{enumerate}
		\textbf{Answer: b) Sets all digital pins 0-7 as outputs with alternating HIGH/LOW} \\
		\textbf{Explanation:} DDRD configures PORTD as output, 0xAA (10101010 binary) sets alternating states.
		
		\item Which material forms the substrate of most PCBs?
		\begin{enumerate}
			\item Aluminum
			\item FR-4
			\item Polycarbonate
			\item Copper
		\end{enumerate}
		\textbf{Answer: b) FR-4} \\
		\textbf{Explanation:} FR-4 glass-reinforced epoxy is the standard PCB base material.
		
		\item What is the primary purpose of decoupling capacitors?
		\begin{enumerate}
			\item Store energy for power outages
			\item Filter high-frequency noise
			\item Increase voltage gain
			\item Limit inrush current
		\end{enumerate}
		\textbf{Answer: b) Filter high-frequency noise} \\
		\textbf{Explanation:} Decoupling capacitors suppress voltage spikes from digital IC switching.
		
		\item Which tool is essential for identifying PCB trace connectivity?
		\begin{enumerate}
			\item Oscilloscope
			\item Multimeter in continuity mode
			\item Logic analyzer
			\item Spectrum analyzer
		\end{enumerate}
		\textbf{Answer: b) Multimeter in continuity mode} \\
		\textbf{Explanation:} Continuity testing verifies electrical connections between points.
		
		\item What does this circuit symbol represent?
		\begin{center}
			\begin{circuitikz}
				\draw (0,0) to[R, l=$R_1$] (2,0);
			\end{circuitikz}
		\end{center}
		\begin{enumerate}
			\item Capacitor
			\item Diode
			\item Resistor
			\item Transistor
		\end{enumerate}
		\textbf{Answer: c) Resistor} \\
		\textbf{Explanation:} Zigzag line is the standard IEEE resistor symbol.
		
		\item How much current can a standard Arduino Uno pin safely source?
		\begin{enumerate}
			\item 10mA
			\item 20mA
			\item 40mA
			\item 100mA
		\end{enumerate}
		\textbf{Answer: b) 20mA} \\
		\textbf{Explanation:} 20mA per pin (40mA absolute max) to prevent damage to microcontroller.
		
		\item Which file format is used for PCB manufacturing files?
		\begin{enumerate}
			\item .INO
			\item .TXT
			\item .GERBER
			\item .JPEG
		\end{enumerate}
		\textbf{Answer: c) .GERBER} \\
		\textbf{Explanation:} Gerber files contain manufacturing data for PCB fabrication.
		
		\item What is the purpose of the Arduino bootloader?
		\begin{enumerate}
			\item To compile sketches
			\item To initialize hardware on startup
			\item To enable programming via USB
			\item To debug code
		\end{enumerate}
		\textbf{Answer: c) To enable programming via USB} \\
		\textbf{Explanation:} Bootloader allows programming without external hardware.
		
		\item Which component protects circuits from reverse polarity?
		\begin{enumerate}
			\item Capacitor
			\item Diode
			\item Resistor
			\item Inductor
		\end{enumerate}
		\textbf{Answer: b) Diode} \\
		\textbf{Explanation:} Diodes allow current flow in only one direction.
		
		\item What does "I2C" stand for?
		\begin{enumerate}
			\item Inter-integrated Circuit
			\item Input/Output Interface Controller
			\item Integrated Internal Communication
			\item Intelligent Interface Chip
		\end{enumerate}
		\textbf{Answer: a) Inter-integrated Circuit} \\
		\textbf{Explanation:} I2C is a serial communication protocol using SDA/SCL lines.
		
		\item The main advantage of surface-mount technology (SMT) over through-hole is:
		\begin{enumerate}
			\item Higher power handling
			\item Easier manual assembly
			\item Smaller component size
			\item Lower cost per component
		\end{enumerate}
		\textbf{Answer: c) Smaller component size} \\
		\textbf{Explanation:} SMT allows higher component density and smaller PCBs.
		
		\item What does this Arduino code snippet do?
		\begin{verbatim}
			analogWrite(9, 128);
		\end{verbatim}
		\begin{enumerate}
			\item Sets pin 9 to 5V output
			\item Reads analog value from pin 9
			\item Generates 50\% duty cycle PWM
			\item Configures pin 9 as input
		\end{enumerate}
		\textbf{Answer: c) Generates 50\% duty cycle PWM} \\
	\textbf{	Explanation:} $128/255 = 50 $ duty cycle (mid-range value).
		
		\item What is the purpose of a ground plane in PCB design?
		\begin{enumerate}
			\item Increase component density
			\item Provide shielding and reduce noise
			\item Improve thermal conductivity
			\item Enhance signal speed
		\end{enumerate}
		\textbf{Answer: b) Provide shielding and reduce noise} \\
		\textbf{Explanation:} Ground planes create low-impedance return paths and reduce EMI.
		
		\item Which tool is used to remove solder from PCB joints?
		\begin{enumerate}
			\item Soldering iron
			\item Desoldering wick
			\item Flux pen
			\item Multimeter
		\end{enumerate}
		\textbf{Answer: b) Desoldering wick} \\
		\textbf{Explanation:} Solder wick (braid) absorbs molten solder for component removal.
		
		\item What is the typical resolution of Arduino Uno's ADC?
		\begin{enumerate}
			\item 8-bit (0-255)
			\item 10-bit (0-1023)
			\item 12-bit (0-4095)
			\item 16-bit (0-65535)
		\end{enumerate}
		\textbf{Answer: b) 10-bit (0-1023)} \\
		\textbf{Explanation:} ATmega328P has 10-bit analog-to-digital converters.
		
		\item What is the purpose of the Arduino `millis()` function?
		\begin{enumerate}
			\item Measure resistance
			\item Generate random numbers
			\item Get elapsed time since power-up
			\item with I2C devices
		\end{enumerate}
		\textbf{Answer: c) Get elapsed time since power-up} \\
		\textbf{Explanation:} `millis()` returns milliseconds since program start for timing operations.
	\end{enumerate}
	
\end{document}