\documentclass{article}
\usepackage{enumitem}
\usepackage{circuitikz}
\usepackage{graphicx}

\begin{document}
	
	\section*{PCB Design Fundamentals \& Tools Quiz}
	
	\subsection*{Core Concepts}
	
	\begin{enumerate}
		\item What is the primary purpose of a schematic diagram in PCB design?
		\begin{enumerate}
			\item To show physical component placement
			\item To illustrate copper trace routing
			\item \textbf{To represent electrical connections abstractly}
			\item To generate manufacturing files
		\end{enumerate}
		\textbf{Explanation:} Schematics show logical connections without physical layout details.
		
		\item Which layer type is NOT typically found in a standard PCB stackup?
		\begin{enumerate}
			\item Copper layer
			\item Solder mask layer
			\item Silkscreen layer
			\item \textbf{Conductive ink layer}
		\end{enumerate}
		\textbf{Explanation:} Standard PCBs use copper conductors, not conductive ink.
		
		\item What is the purpose of solder mask on a PCB?
		\begin{enumerate}
			\item To increase conductivity
			\item \textbf{To prevent solder bridges and protect copper}
			\item To label components
			\item To improve thermal dissipation
		\end{enumerate}
		\textbf{Explanation:} Solder mask insulates copper traces and prevents accidental connections.
		
		\item What does "DRC" stand for in PCB design tools?
		\begin{enumerate}
			\item Digital Resistance Calculator
			\item \textbf{Design Rule Check}
			\item Diagrammatic Routing Configuration
			\item Designated Routing Channel
		\end{enumerate}
		\textbf{Explanation:} DRC verifies design against manufacturing constraints.
		
		\item Which via type connects only specific layers in a multilayer PCB?
		\begin{enumerate}
			\item Through-hole via
			\item \textbf{Blind via}
			\item Microvia
			\item Buried via
		\end{enumerate}
		\textbf{Explanation:} Blind vias connect surface layers to internal layers without going through the entire board.
		
		\item What is the purpose of keep-out zones in PCB layout?
		\begin{enumerate}
			\item To mark component-free areas for cooling
			\item To designate high-voltage isolation areas
			\item To reserve space for mechanical constraints
			\item \textbf{All of the above}
		\end{enumerate}
		\textbf{Explanation:} Keep-out zones restrict placement/routing for various design constraints.
		
		\item Which factor is MOST critical when determining trace width?
		\begin{enumerate}
			\item Color scheme
			\item \textbf{Current carrying capacity}
			\item Signal frequency
			\item Component density
		\end{enumerate}
		\textbf{Explanation:} Trace width primarily determines current handling capability before overheating.
		
		\item What is the purpose of a thermal relief in a PCB pad?
		\begin{enumerate}
			\item To increase current capacity
			\item \textbf{To ease soldering by reducing heat dissipation}
			\item To improve RF performance
			\item To enhance mechanical strength
		\end{enumerate}
		\textbf{Explanation:} Thermal reliefs reduce heat transfer during soldering while maintaining electrical connection.
	\end{enumerate}
	
	\subsection*{Design Tools \& Workflow}
	
	\begin{enumerate}[resume]
		\item Which file format is the industry standard for PCB manufacturing data?
		\begin{enumerate}
			\item .PDF
			\item .DXF
			\item \textbf{.GERBER}
			\item .STEP
		\end{enumerate}
		\textbf{Explanation:} Gerber files (RS-274X) contain layer-by-layer fabrication data.
		
		\item What is the purpose of a Bill of Materials (BOM) in PCB design?
		\begin{enumerate}
			\item To show component placement
			\item \textbf{To list all components with specifications}
			\item To document routing topology
			\item To specify test points
		\end{enumerate}
		\textbf{Explanation:} BOM provides comprehensive component list for procurement and assembly.
		
		\item In PCB design software, what is "forward annotation"?
		\begin{enumerate}
			\item Adding comments to schematics
			\item \textbf{Pushing schematic changes to PCB layout}
			\item Generating 3D models
			\item Creating assembly drawings
		\end{enumerate}
		\textbf{Explanation:} Forward annotation propagates schematic changes to the PCB editor.
		
		\item Which tool is essential for verifying net connectivity between schematic and PCB?
		\begin{enumerate}
			\item DRC Checker
			\item \textbf{ERC (Electrical Rule Check)}
			\item LVS (Layout vs. Schematic)
			\item Signal Integrity Analyzer
		\end{enumerate}
		\textbf{Explanation:} LVS tools verify layout matches schematic connectivity.
		
		\item What does "SMT" stand for in PCB assembly?
		\begin{enumerate}
			\item Standardized Mounting Technology
			\item \textbf{Surface Mount Technology}
			\item System Module Termination
			\item Soldering Matrix Technique
		\end{enumerate}
		\textbf{Explanation:} SMT refers to surface-mount components soldered directly to PCB pads.
		
		\item Which design element helps mitigate electromagnetic interference (EMI)?
		\begin{enumerate}
			\item Decorative silkscreen patterns
			\item \textbf{Proper ground planes and shielding}
			\item Increased trace spacing
			\item Thermal vias under components
		\end{enumerate}
		\textbf{Explanation:} Continuous ground planes and shields contain electromagnetic fields.
		
		\item What is the purpose of a fiducial marker on a PCB?
		\begin{enumerate}
			\item Decorative branding
			\item \textbf{Automated optical alignment for assembly}
			\item Electrical test point
			\item Heat dissipation point
		\end{enumerate}
		\textbf{Explanation:} Fiducials provide reference points for pick-and-place machines.
		
		\item Which feature allows high-density component placement in modern designs?
		\begin{enumerate}
			\item Through-hole components
			\item \textbf{BGA (Ball Grid Array) packages}
			\item DIP sockets
			\item Wire-wrapping
		\end{enumerate}
		\textbf{Explanation:} BGAs allow hundreds of connections under the package.
	\end{enumerate}
	
	\subsection*{Design Rules \& Manufacturing}
	
	\begin{enumerate}[resume]
		\item What is the typical copper thickness for standard PCBs?
		\begin{enumerate}
			\item 0.1mm
			\item \textbf{$1oz/ft^2$ (35  micro meter)}
			\item 0.5mm
			\item 2mm
		\end{enumerate}
		\textbf{Explanation:} 1oz/ft² is the standard baseline copper weight.
		
		\item What clearance is typically required between traces for 120V operation?
		\begin{enumerate}
			\item 0.1mm
			\item 0.2mm
			\item \textbf{0.5mm}
			\item 1mm
		\end{enumerate}
		\textbf{Explanation:} IPC-2221 requires 0.5mm minimum for 120V signals.
		
		\item Which manufacturing defect is caused by insufficient solder paste?
		\begin{enumerate}
			\item Tombstoning
			\item \textbf{Solder starvation}
			\item Solder bridging
			\item Cold joint
		\end{enumerate}
		\textbf{Explanation:} Insufficient paste leads to poor solder joints.
		
		\item What does "HASL" stand for in PCB finishing?
		\begin{enumerate}
			\item Heat Activated Solder Layer
			\item \textbf{Hot Air Solder Leveling}
			\item High Accuracy Surface Layer
			\item Hybrid Assembly Solder Lead
		\end{enumerate}
		\textbf{Explanation:} HASL is a common surface finish where molten solder is applied and leveled.
		
		\item Which design consideration is CRITICAL for high-frequency signals?
		\begin{enumerate}
			\item Component color coding
			\item \textbf{Controlled impedance routing}
			\item Silkscreen legibility
			\item Through-hole pad sizes
		\end{enumerate}
		\textbf{Explanation:} Impedance control prevents signal reflections in high-speed circuits.
		
		\item What is the purpose of a mouse bite in PCB panelization?
		\begin{enumerate}
			\item To reduce EMI
			\item \textbf{To facilitate board separation}
			\item To improve heat dissipation
			\item To increase rigidity
		\end{enumerate}
		\textbf{Explanation:} Mouse bites are perforations allowing easy depanelization.
		
		\item Which software is NOT a professional PCB design tool?
		\begin{enumerate}
			\item Altium Designer
			\item KiCad
			\item Cadence Allegro
			\item \textbf{AutoCAD LT}
		\end{enumerate}
		\textbf{Explanation:} AutoCAD LT is general CAD, not specialized for PCB design.
		
		\item What does "IPC" stand for in PCB standards?
		\begin{enumerate}
			\item International Processing Council
			\item \textbf{Institute for Printed Circuits}
			\item Integrated Production Criteria
			\item Interconnected Package Components
		\end{enumerate}
		\textbf{Explanation:} IPC develops standards for PCB design and manufacturing.
		
		\item Which layer contains component outlines and reference designators?
		\begin{enumerate}
			\item Copper layer
			\item Solder mask layer
			\item \textbf{Silkscreen layer}
			\item Paste mask layer
		\end{enumerate}
		\textbf{Explanation:} Silkscreen provides component identification and outlines.
		
		\item What is the purpose of a via tenting?
		\begin{enumerate}
			\item To increase current capacity
			\item \textbf{To cover vias with solder mask}
			\item To connect multiple layers
			\item To improve thermal conduction
		\end{enumerate}
		\textbf{Explanation:} Via tenting prevents solder wicking and contamination.
	\end{enumerate}
	
	\subsection*{Tool-Specific Features}
	
	\begin{enumerate}[resume]
		\item In KiCad, what is the purpose of the "CvPCB" tool?
		\begin{enumerate}
			\item Circuit simulation
			\item \textbf{Component footprint assignment}
			\ copper pour creation
			\item 3D model generation
		\end{enumerate}
		\textbf{Explanation:} CvPCB links schematic symbols to physical footprints.
		
		\item What is the primary function of a "room" in Altium Designer?
		\begin{enumerate}
			\item 3D visualization space
			\item \textbf{Grouping components for placement}
			\item Defining keep-out areas
			\item Creating power planes
		\end{enumerate}
		\textbf{Explanation:} Rooms allow collective component placement and replication.
		
		\item Which Eagle CAD feature allows parametric design?
		\begin{enumerate}
			\item BRD files
			\item \textbf{User Language Programs (ULPs)}
			\item Layer sets
			\item Design blocks
		\end{enumerate}
		\textbf{Explanation:} ULPs enable scripting and parametric operations.
		
		\item What is the purpose of "differential pair routing" in high-speed design tools?
		\begin{enumerate}
			\item To reduce component count
			\item \textbf{To maintain constant impedance for matched-length traces}
			\item To increase current capacity
			\item To simplify schematics
		\end{enumerate}
		\textbf{Explanation:} Differential pairs require precise length matching for signal integrity.
		
		\item Which Altium feature automates component placement based on schematics?
		\begin{enumerate}
			\item ActiveRoute
			\item \textbf{Component placement panel}
			\item Draftsman
			\item Output job files
		\end{enumerate}
		\textbf{Explanation:} Placement tools suggest optimal component arrangements.
	\end{enumerate}
	
	\subsection*{Advanced Considerations}
	
	\begin{enumerate}[resume]
		\item What is the purpose of a stitching via in PCB design?
		\begin{enumerate}
			\item To increase mechanical strength
			\item \textbf{To connect ground planes across layers}
			\item To enhance thermal dissipation
			\item To reduce EMI emission
		\end{enumerate}
		\textbf{Explanation:} Stitching vias maintain low-impedance ground connections.
		
		\item Which design technique reduces crosstalk in dense layouts?
		\begin{enumerate}
			\item Increasing trace thickness
			\item \textbf{Guard traces with ground vias}
			\item Using right-angle traces
			\item Maximizing parallel routing
		\end{enumerate}
		\textbf{Explanation:} Guard traces isolate sensitive signals from interference.
		
		\item What does "HDI" stand for in modern PCB technology?
		\begin{enumerate}
			\item High-Density Integration
			\item \textbf{High-Density Interconnect}
			\item Hybrid Digital Interface
			\item High Definition Interposer
		\end{enumerate}
		\textbf{Explanation:} HDI PCBs use microvias and thin materials for dense routing.
		
		\item Which simulation is ESSENTIAL for power delivery networks?
		\begin{enumerate}
			\item Thermal stress analysis
			\item \textbf{DC voltage drop analysis}
			\item RF radiation pattern
			\item Acoustic resonance
		\end{enumerate}
		\textbf{Explanation:} Voltage drop analysis ensures adequate power to all components.
		
		\item What is the primary benefit of embedded components in PCBs?
		\begin{enumerate}
			\item Reduced cost
			\item \textbf{Improved signal integrity and miniaturization}
			\item Easier rework
			\item Simplified manufacturing
		\end{enumerate}
		\textbf{Explanation:} Embedded components save space and reduce parasitic effects.
	\end{enumerate}
	
\end{document}
