%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%       Syllabus: Communication Engineering I
%       Course No: EEE 3117
%       Rajshahi University of Engineering & Technology
%       (Fully Enumerated Version)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[a4paper,12pt]{article}

\usepackage{graphicx} % Required for inserting images
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{subcaption}
% -----------------------
% Package Imports
% -----------------------

% Set page margins
\usepackage[a4paper, top=1in, bottom=0.8in, left=1.1in, right=0.8in]{geometry}

% Use Times New Roman font
\usepackage{times}

% Add page numbering
\pagestyle{plain}
\usepackage{multirow}
% Enable graphics inclusion
\usepackage{graphicx}
\usepackage{float}
% Enable code listings
\usepackage{listings}
\usepackage{xcolor} % For customizing code colors

\usepackage{enumitem}

\title{
	\Huge Syllabus: Communication Engineering I \\
	\vspace{0.2cm}
	\large Course No: EEE 3117
}
\author{
	Department of Electrical and Electronic Engineering \\
	Rajshahi University of Engineering \& Technology
}
\date{}


\begin{document}
	
	\maketitle
	\thispagestyle{empty}
	
	\section*{Course Content}
	\begin{enumerate}[leftmargin=*, align=left]
		
		\item \textbf{Overview of Communication Systems}
		\begin{enumerate}[label*=\arabic*.]
			\item Basic principles, Fundamental elements, Message sources, Input-output transducers.
			\item Transmission policies.
			\item Analog and digital communication systems, their advantages and disadvantages.
			\item Introduction to communication networks.
			\item Communication traffics.
			\item Telephone communication.
			\item Satellite communication.
			\item RADAR communication.
			\item Fiber optic communication.
			\item Cellular communication.
			\item Radio and TV broadcasting.
			\item Radio switching systems.
			\item Cognitive radio networks etc.
		\end{enumerate}
		
		\item \textbf{Transmitter}
		\begin{enumerate}[label*=\arabic*.]
			\item Elements of transmitter and their functions.
			\item Concept of wireless transmission with antennas.
			\item Transmit diversity schemes.
			\item Channel state information at the transmitter.
			\item Transmission delay.
			\item \textbf{Digitization:}
			\begin{enumerate}[label*=\arabic*.]
				\item Sampling and its classification, Sampling theorem, Nyquist criterion, Aliasing effect and its elimination.
				\item Quantizing, Quantization noise, Non-uniform quantization.
				\item Signal to quantization error Ratio.
				\item Encoding, Line coding formats.
				\item Sample and hold circuit.
				\item Quantizer and encoder circuits.
			\end{enumerate}
			\item \textbf{Modulation:}
			\begin{enumerate}[label*=\arabic*.]
				\item Modulation and its principle, Importance of modulation.
				\item Definition, Graphical representation, Generation.
				\item Modulated signal’s expression.
				\item Frequency spectrum and bandwidth requirements of AM, FM and PM.
				\item Design and fabrication of AM, FM and PM transmitter circuits.
			\end{enumerate}
		\end{enumerate}
		
		\item \textbf{Transmission Media}
		\begin{enumerate}[label*=\arabic*.]
			\item Types of media and their properties.
			\item Principle of information transmission through wire, Coaxial cable, Waveguide, Optical fiber, Radio link etc.
			\item Channel and its classification.
			\item PDF, CDF and MGF of channel gain.
			\item Modeling of fading channels.
			\item Expressions of received signals.
			\item Propagation delay.
			\item Bandwidth.
			\item Channel capacity and its classification.
			\item Outage probability.
			\item Bit error rate, Symbol error rate and their calculations.
		\end{enumerate}
		
		\item \textbf{Noise}
		\begin{enumerate}[label*=\arabic*.]
			\item Effects of noise on the transmission.
			\item Sources of noise.
			\item Characteristics of various types of noise.
			\item Signal-to-noise ratio.
			\item Noise figure.
		\end{enumerate}
		
		\item \textbf{Fading}
		\begin{enumerate}[label*=\arabic*.]
			\item Types of fading.
			\item Measures of fading channels.
		\end{enumerate}
		
		\item \textbf{Interference}
		\begin{enumerate}[label*=\arabic*.]
			\item Types of interference and their mitigation techniques.
		\end{enumerate}
		
		\item \textbf{Correlation}
		\begin{enumerate}[label*=\arabic*.]
			\item Types of correlation and their mitigation technologies.
		\end{enumerate}
		\newpage
		\item \textbf{Digital Link}
		\begin{enumerate}[label*=\arabic*.]
			\item Design of digital link.
		\end{enumerate}
		
		\item \textbf{Receiver}
		\begin{enumerate}[label*=\arabic*.]
			\item Receivers, Elements of receiver and their functions.
			\item Superheterodyne receiver.
			\item Design and fabrication of matched filter circuit.
			\item Analog, Digital and correlator receiver circuits.
			\item Receiving antennas.
			\item Channel state information at the receiver.
		\end{enumerate}
		
		\item \textbf{Demodulation and Decoding}
		\begin{enumerate}[label*=\arabic*.]
			\item Principle of demodulation and decoding.
			\item Importance of demodulation.
			\item Demodulation and decoding of AM, FM and PM modulated signals.
			\item Design and fabrication of AM, FM and PM receiver circuits.
		\end{enumerate}
		
	\end{enumerate}
	
\end{document}